---
title: "Variable Report"
author: "Tom Detzel"
subtitle: Cuny MSDA 624 Project 2
date: "11/25/2017"
output:
    prettydoc::html_pretty:
    theme: leonid
    highlight: github
    toc: yes
---

```{r setup, warning=F, message=F, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

# load required packages
suppressMessages(library(easypackages))
suppressMessages(libraries("tidyverse", "nnet", "kernlab", "caret", "car", "earth", "pander", "randomForest", "mlbench","ModelMetrics", "lars", "MASS", "stats", "pls", "psych", "corrplot", "Hmisc", "mi", "betareg", "mice", "rpart", "party", "partykit", "gbm", "ipred", "VIM"))

```

### Get the data

```{r warning=F, message=F}

# read in the data
df <- read.csv("data/StudentData.csv", header=T, strip.white=T)
# str(df)

```

### Get my variables

```{r}
df <- df[, 25:33]

```

### Missingness  

Missing values aren't a big problem with these vars.

```{r}
## assumes '0' is an NA
zero_vals <- data.frame(cbind(colSums(df==0)))
colnames(zero_vals) <- "zero_count"
pander(zero_vals)

```

### Work with complete cases  
  
We lost 40 observations, a de minimus number. 

```{r}

# recode '0' as NA
df[df==0] <- NA

# complete cases
df_c <- na.omit(df) 

#reorder cols to make PH last
df_c <- df_c[c(1,3:9,2)]

```

### Exploratory plots

Histograms and density plots show the bimodal character of several variables: Bowl.Setpoint, Pressure.Setpoint, Alch.Rel, Carb.Rel, and Balling.Lvl.

### Histograms

```{r fig.height=12, fig.width=12}

# histograms for each variable
par(mfrow=c(4,2))
for(i in c(1:8)) {
  hist(df_c[,i], main=names(df_c)[i], col="lightblue")
}

```

### Density plots

```{r fig.height=10, fig.width=12}

# density plot for each var
par(mfrow=c(2,4))
for(i in 1:8) {
  plot(density(df_c[,i]), main=names(df_c)[i])
}

```

### Boxplots

```{r fig.height=10, fig.width=12}
# boxplots
par(mfrow=c(2,4))
for(i in 1:8) {
  boxplot(df_c[,i], main=names(df_c)[i], col="lightblue")
}
```

### XY plots

Pressure.Vacuum, Oxygen Filler, Bowl.Setpoint and Carb.Rel all have a postive association with PH; Pressure.Setpoint has a modest negative correlation.

```{r}

par(mfrow=c(2,4))
for(i in 1:8) {
  plot(df_c[,i], df_c$PH, xlab=colnames(df_c)[i], ylab="PH", col="lightblue")
  abline(lsfit(df_c[,i], df_c$PH), col="red", lwd=2)
}
```

### Correlation plot
  
Alch.Rel, Carb.Rel and Balling.Lvl are strongly correlated to each other. None of the vars is strongly correlated with PH.

```{r}

# correlation plot 
correlations <- cor(df_c)
corrplot(correlations, method="circle")

```

### Variable importance

We'll fit linear, Random Forest and Cubist models to test variable importance. Results suggest that Air.Pressure isn't an important variable and could be eliminated to simplify the model.
  
```{r}
# fit models

# Run lm using 10-fold cross-validation
trainControl <- trainControl(method="repeatedcv", number=10, repeats=1)
metric <- "RMSE"

# LM
set.seed(100)
fit.lm <- train(PH~., data=df_c, method="lm", metric=metric,
                preProc=c("center", "scale"), trControl=trainControl)

# Cubist
set.seed(100)
fit.cubist <- train(PH~., data=df_c, method="cubist", metric=metric,
                    ppreProc=c("center", "scale"), trControl=trainControl)

# Random Forest
set.seed(100)
fit.rf <- train(PH~., data=df_c, method="rf", metric=metric,
                preProc=c("center", "scale"),
                trControl=trainControl,
                importance=TRUE)

```

### Error comparison

```{r}

modelResults <- resamples(list(LM=fit.lm, Cubist=fit.cubist, RF=fit.rf))

summary(modelResults)

```

### Varible importance, Cubist

```{r}

varImp(fit.cubist)

```

### Variable importance plot, Random Forest

```{r}

varImp(fit.rf)

```

### Variable importance, linear fit

Results show that Pressure.Vaccum, Bowl.Setpoint, Pressure.Setpoint, Alch.Rel and Balling.Lvl all are significant at p <.001. In this model, these predictors explain ony 19 percent of the variation in PH.

```{r}
summary(fit.lm)
```

